---
title: "A3: SPL Library Checkouts"
output: html_document
---
```{r, echo = FALSE}
jg_df <- read.csv("John_Green_Checkouts.csv", stringsAsFactors = FALSE)
```

### Introduction
Using the Seattle Public Library data, I chose to analyze the books by John Green that were checked out in Seattle. I chose this author because he was very popular when I was younger, and I wanted to understand how popular his books have been recently. I was also curious about whether certain factors related to the level of popularity of his books. This leads me to my main research question:

- *How popular has John Green's books been from 2006 to 2022?*

In order to do so, I will focus on the **checkout year**, **number of checkouts**, **subjects of the book**, **publication year**, and **publishing company** when calculating summary statistics and creating data visualizations. I will use these variables to analyze the following trends:

- *How have the most checked out books written by John Green changed overtime?*
- *How does the number of John Green's books that were published correlate to the number of John Green's books that were checked out overtime?*
- *How has the popularity of John Green's books' genres changed overtime?*

### Summary Information
```{r, echo = FALSE, warning = FALSE, message = FALSE, code = readLines("summary.R")}
```
In order to assess John Green's popularity throughout the years and factors that may be associated with his popularity, I calculated summary statistics. First, I found that the book written by John Green with the most number of checkouts in Seattle from 2006 to 2022 was `r book_most_checkouts`. This helps me understand John Green's most popular book. Then, I calculated the most popular genre that John Green writes for by calculating the number of checkouts per genre. The genre with the most checkouts is `genre_most_checkouts`. Regarding the time period when John Green's books were most popular, I calculated the month and year when the most John Green books were checked out, which is `date_most_checkouts`. As for the publishing company, I found that the most checkouts of books written by John Green were published by `publish_cmpny_most_checkout`. Finally, most of John Green's book checkouts were checked out as an `most_medium`.

### The Dataset

- Who collected/published the data?  
- What are the parameters of the data (dates, number of checkouts, kinds of books, etc.)?
- How was the data collected or generated?  
- Why was the data collected?  
- What, if any, ethical questions do you need to consider when working with this data?  
- What are possible limitations or problems with this data?   (at least 200 words)


### First Trends Over Time Chart

Include a chart. Make sure to describe why you included the chart, and what patterns emerged

The first chart that you will create and include will show the trend over time of your variable/topic/interest. Think carefully about what you want to communicate to your user (you may have to find relevant trends in the dataset first!). Here are some requirements to help guide your design:

- Show more than one, but fewer than ~10 trends
  - For example, two books, or 5 different mediums
- You must have clear x and y axis labels
- The chart needs a clear title 
- You need a legend for your different line colors and a clear - legend title
In your .Rmd file, make sure to describe why you included the chart, and what patterns emerged

When we say "clear" or "human readable" titles and labels, that means that you should not just display the variable name.

Here's an example of how to run an R script inside an RMarkdown file:

``````{r, echo = FALSE, warning = FALSE, code = readLines("chart1_example.R")}
```

### Second Trends Over Time Chart

Include a chart. Make sure to describe why you included the chart, and what patterns emerged

The second chart that you will create and include will show another trend over time of your variable/topic/interest. Think carefully about what you want to communicate to your user (you may have to find relevant trends in the dataset first!). Here are some requirements to help guide your design:

- Show more than one, but fewer than ~10 trends
  - For example, two books, or 5 different mediums
- You must have clear x and y axis labels
- The chart needs a clear title 
- You need a legend for your different line colors and a clear - legend title
In your .Rmd file, make sure to describe why you included the chart, and what patterns emerged

When we say "clear" or "human readable" titles and labels, that means that you should not just display the variable name.

Here's an example of how to run an R script inside an RMarkdown file:

``````{r, echo = FALSE, code = readLines("chart2_example.R")}
```

### Your Choice

The last chart is up to you. It could be a line plot, scatter plot, histogram, bar plot, stacked bar plot, and more. Here are some requirements to help guide your design:

- You must have clear x and y axis labels
- The chart needs a clear title 
- You need a legend for your different line colors and a clear legend title

Here's an example of how to run an R script inside an RMarkdown file:

``````{r, echo = FALSE, code = readLines("chart2_example.R")}
```


