---
title: 'A3: SPL Library Checkouts'
output:
  html_document: default
  pdf_document: default
---
```{r, echo = FALSE}
jg_df <- read.csv("ALL_JG_BOOKS.csv", stringsAsFactors = FALSE)
```

### Introduction
Using the Seattle Public Library data, I chose to analyze the books by John Green that were checked out in Seattle. I chose this author because he was very popular when I was younger, and I wanted to understand how popular his books have been recently. I was also curious about whether certain factors related to the level of popularity of his books. This leads me to my main research question:

- *How popular has John Green's books been from 2007 to 2023?*

In order to do so, I will focus on the **Checkouts**, **Title**, **CheckoutYear**, **CheckoutMonth**, **Publisher**, and **MaterialType** when calculating summary statistics and creating data visualizations. I will not use the "Subjects" column because many of John Green's books fall under the same genres, and any unique genres tend to fall only under one book title. I will not be using the "PublicationYear" column due to the large variation in format. Using my selected variables, I will analyze the following trends:

- *How have the most checked out books written by John Green changed overtime?*
- *How have the most common materials checked out for John Green's books changed overtime?*
- *How has the checkouts of "The Fault in Our Stars" changed over time?*

### Summary Information
```{r, echo = FALSE, warning = FALSE, message = FALSE, code = readLines("summary.R")}
```
In order to assess John Green's popularity throughout the years and factors that may be associated with his popularity, I calculated summary statistics that use the variables specified in the introduction. First, I wanted to know which piece of work written by John Green was the most popular, which I found by pulling the title of the item with the most amount of checkouts. I found that the item in the Seattle Public Library written by John Green with the most number of checkouts was "`r item_most_checkouts`". For this statistic, I purposefully did not combine items with a similar book title because these differences in title formats are due to differences in the item's material type which can potentially impact how often an item is checked out. (For instance, John Green's book called "An Abundance of Katherines" can also be found in the dataframe under the title "An Abundance of Katherines (Unabridged)" because the two are different types of mediums.) 

Then, I wanted to understand if other factors impacted John Green's popularity in the Seattle Public Library. I did so by calculating how often an item was checked out based on factors such as the date of the check out and the medium of the item. First, to assess the time period when John Green's books were most popular, I calculated the month and year when the most John Green books were checked out to be `r date_most_checkouts`. To see if annual holidays or events impacted the popularity of John Green's books, I calculated the month (regardless of year) with the most amount of checkouts. I found this month to be `r month_highest_checkouts`.

As for type of material, John Green's book checkouts were most checked out as a `r medium_most_checkouts`. Lastly, in order to determine if the material type changed the popularity of a story, I found the *book* with the most amount of checkouts, which was also "`r book_most_checkouts`". This tells me that the material did not impact the most popular story. When calculating this last statistic, I did change the format of the entry in the "Title" column to match the same story because all of the entries were the same type of medium. (For instance, the story "Looking for Alaska" was entered as "Looking for Alaska / John Green." and "Looking for Alaska / a novel by John Green." although both are books. These entries do have different publishers, but because publishers don't change the content of a story, I did not think the difference in publishers was relevant and grouped these titles under the same title.)

### The Dataset

> Who collected/published the data?  

- Before 2017, George Legrady collected the checkout data since 2005 to use for his piece of artwork called "Making the Invisible Visible". The Seattle Public Library then began data collection in 2017 and combined their data with Legrady's data in order to publish everything to the Open Data program.
- My custom dataframe uses this data, but it is filtered to only include checkout items written by the American author John Green.
  
> What are the parameters of the data (dates, number of checkouts, kinds of books, etc.)?

- The parameters of the data are listed below:
  - UsageClass (whether the check out was digital or physical)
  - CheckoutType (the vendor tool used for the check out)
  - MaterialType (the medium of the check out)
  - CheckoutYear
  - CheckoutMonth    
  - Checkouts (number of check outs of the item)
  - Title (of the item checked out)
  - ISBN
  - Creator
  - Subjects (a list of genres of the item)
  - Publisher
  - PublicationYear
  
> How was the data collected or generated?  

- The digital checkout data was collected through the Seattle Public Library's vendors (who are Overdrive, hoopla, Freegal, and RBDigital). Data on physical checkouts was collected from the Legrady artwork data archives from 2005 to 2016 and from the Horizon ILS since 2016.

> Why was the data collected?  

- According to Seattle's Open Data website, data is published by the city's departments and put into their Open Data program. This Open Data program was created so that the data collected by Seattle was easily accessible to the public. They wanted the availability of the data to increase Seattle's quality of life and promote a supportive and beneficial community to its residents. The program wanted the data to be used towards its goal of constant self-improvement.
- Before this, George Legrady collected the checkout data to create his artwork that displays summarized and recent data of the library's checkouts.

> What, if any, ethical questions do you need to consider when working with this data? 

- One ethical question to consider is how the Seattle Public Library alerted its users that their data was being collected. The [article](https://shelftalkblog.wordpress.com/2022/12/06/how-a-digital-artwork-helped-the-seattle-public-library-lead-in-book-data/) that explains how the SPL checkout data came to be claims that Legrady's data did not collect any information on its users but does not specify further. Although the data remains anonymous, users of the library may not be comfortable with any of their check out information being collected as the items' subjects can read into the users' personal interests or political beliefs. Considering how many websites force a user to accept their terms and conditions in order to use the website, the library may follow a similar policy for users to check out items. This means that for users who require the use of the library's items, they would be forced to share their data in order to access the library's resources.

> What are possible limitations or problems with this data?

- One limitation of the data is that it does not record renewal check outs and only records the initial check out. This means that the data is likely to have underestimated the number of total check outs per month as it is common for users to extend their time limit. 

- Another problem of the data is that it is very messy due to its variation in sources. The different sources used to collect the data means that different types of information may have been collected using different methods. For example, the source for the physical checkouts changed during 2016, so some entries have blank fields meaning the data is inconsistent depending on what time frame is being viewed. Furthermore, the original dataset is messy as some of the entries use unreadable characters and the format of data entries differs. This can be seen when looking at the "PublicationYear" column which can make filtering for a certain data value difficult. This also includes book titles that are written in English or Spanish or written with different capitalized letters (such as "unabridged" versus "Unabridged").

- A third problem relates to the Subjects column. The subject of each entry is a string of subjects. This means that each subject entry does not pertain to one unique subject. For example, when looking at the subjects for the book titled "An Abundance of Katherines" and comparing it to "Paper Towns", the entire subject entry itself is different. However, after taking a closer look at each subject in the list, we are able to find that both books share Young Adult Fiction as a subject. 

- Lastly, in my filtered dataset, the creator is filtered to "John Green", but this doesn't mean that all creators are the same person. This means that creators who share the same name are included in this dataset, so I need to further clean the data to only include items created by the American author John Green. 

- In order to limit these issues in my custom dataframe, I applied multiple filters to the "Creator" column in order to obtain as many rows as possible that include the name "John Green". Then, I went through the "Title" column in alphabetical order to check if each item was written by the American author John Green or written in a different language. I omitted any data entries that were not written in English or written by a different author. As explained in the summary information section, I did not combine the same works together (such as combining all works of "The Fault in Our Stars" because the different in title entries were due to a difference in material type. Lastly, I did not use the ISBN column when using my custom dataset because it contained a significantly large amount of "NA" values. 

### First Trends Over Time Chart
```{r, echo = FALSE, warning = FALSE, message = FALSE, code = readLines("chart1.R")}
```

I included a chart that shows the trends of the number of checkouts of the English ebooks written by John Green between 2008 and 2023. I wanted to visualize how popular John Green's books were overtime, so I decided to do so by analyzing the number of checkouts for each ebook by month and year. I chose to only visualize ebooks because this material was more common than the audiobooks, and there were only ebooks and audiobooks in list of materials. I also only chose the English versions because I didn't want to combine the same titles in English and Spanish in case language plays a factor in popularity. The chart reveals that "The Fault in Our Stars" had the greatest peak in checkouts around 2014. Although not as great, "Turtles All the Way Down" also experienced significant peak in checkouts from 2016 to 2020. The annual total number of checkouts for John Green's English ebooks has never exceeded 250 checkouts. Also, seeing as how the peaks have declined overtime, John Green's English ebooks don't seem to be as popular as they were in the early to mid 2010's. 

### Second Trends Over Time Chart
```{r, echo = FALSE, warning = FALSE, message = FALSE, code = readLines("chart2.R")}
```

This line plot shows how the material of the checkout (for works by John Green) have changed over time. I wanted to understand the popularity of each material type throughout the years as the material of a book is a potential factor to an author's popularity. For instance, if an author publishes more works as a physical book when more readers prefer ebooks, then the author's popularity could decrease. This line plot shows that the number of ebook Checkouts has been greater than the number of audiobook checkouts on average for John Green's works. However, it also shows the general decline in ebook checkouts and the rise in audiobook checkouts. This suggests that more customers are interested in audiobooks, so John Green may want to consider in publishing more audiobook stories to increase his popularity and success. 

### Stacked Bar Plot
```{r, echo = FALSE, warning = FALSE, message = FALSE, code = readLines("chart3.R")}
```

This bar plot shows the annual checkout totals for "The Fault in Our Stars". From the first chart, "The Fault in Our Stars" seemed to be when John Green's popularity dramatically peaked, so I wanted to further analyze the checkout trends of this story specifically. The bar plot includes both the English and Spanish versions of the book because I wanted to analyze the book's popularity regardless of language. I also created stacked bars based on material type because I wanted to see how the proportion of audiobook checkouts to ebook checkouts changed over time. It's also important to note that the English and Spanish audiobooks were published after the English ebook version was published which explains why there are no audiobook checkouts from 2013 to 2016.

The bar plot reveals that specifically for "The Fault in Our Stars", the book's popularity peaked at more than 2000 checkouts in 2014 but decreased significantly afterwards. On average, the checkouts for ebooks are much greater than the checkouts for audiobooks. However, the proportion of audiobook checkouts has increased from 2017 to January 2023. This can be seen in the 2022 bar in which the audiobook checkouts make up half of the total checkouts for 2022. 



